---
title: "Are Mass Shooting Really That Random?"
authors: Emily Tumbaco, Lorena Mancino, Rob Golebieski
output: html_notebook
---

# Problem 1 

## Importing the Dataset & Installing Necessary Packages
```{r}
shootings.dataset<-read.csv("data.csv") #importing the dataset 
shootings.dataset
```

```{r}
# install.packages('dplyr') # this package contains the mutate function which we used below 

library(dplyr)

shootings.dataset <- mutate(shootings.dataset, date= as.Date(date, format = "%m/%d/%Y"))
shootings.dataset$date # converting the dates in the data set to actual dates in R
```


## Graph 1: Comparison of Mass Shootings Fatalities Before 9/11/2001 and After 9/11/2001

```{r}
a <- shootings.dataset[86:118, 3]

b <- shootings.dataset[86:118, 5]

x <- shootings.dataset[1:85, 3]

y <- shootings.dataset[1:85, 5]


par(mfrow=c(2,2))
plot(a, b, main = "Mass Shootings Fatalities Before 9/11/2001",xlab = "Year", ylab = "Fatalities")
plot(x, y,main = "Mass Shootings Fatalities After 9/11/2001", xlab = "Year", ylab = "Fatalities")
```


# Problem 4 

## 2019 Raster Plot

```{r}
from.date <- as.Date('2019-01-01', format = "%Y-%m-%d")
to.date <- as.Date('2019-12-31', format = "%Y-%m-%d")

plot(NA, xlim=c(from.date, to.date), main="2019 Raster Plot", ylim=c(0,10), xlab="Day of Year", ylab="y")

segments(shootings.dataset$date, 0, y1=10)
```


## 2018 Raster Plot

```{r}
from.date.2 <- as.Date('2018-01-01', format = "%Y-%m-%d")
to.date.2 <- as.Date('2018-12-31', format = "%Y-%m-%d")

plot(NA, xlim=c(from.date.2, to.date.2), ylim=c(0,10), main="2018 Raster Plot",xlab="Day of Year", ylab="y")

segments(shootings.dataset$date, 0, y1=10)
```


## 1999 Raster Plot

```{r}
from.date.1 <- as.Date('1999-01-01', format = "%Y-%m-%d")
to.date.1 <- as.Date('1999-12-31', format = "%Y-%m-%d")

plot(NA, xlim=c(from.date.1, to.date.1), ylim=c(0,10), main="1999 Raster Plot",xlab="Day of Year", ylab="y")

segments(shootings.dataset$date, 0, y1=10)
```
From looking at the raster plots, the mass shootings seem to occur in clusters. Every 3 months, there are approximately 3 shootings clustered together. The highest frequency of shootings occurs in the Fall months, and the lowest frequency occurs in the Summer. 


# Problem 6

2019 Results 

## Calculating One Coefficient of Variation For The Year 2019

```{r}
tau.2019 <- as.numeric(-diff(shootings.dataset$date[2:11])) #Calculating the time between two successive mass shootings in the year 2019

tau.sd<- sd(tau.2019) 
tau.mean <- mean(tau.2019)

coef.variation <- tau.sd/tau.mean # Calculating the coefficient of variation for 2019

coef.variation-1
```


## Bootstrap Function with 2019 Results 

```{r}
#' Simulates the bootstrap coefficient variation from a case where the shootings are happening at random for a given year
#'
#' @param days number of days in a year 
#' @param shootings number of shooting that given year 
#'
#' @return coefficient variation for that year 


bootstrap.coef.var <- function(days, shootings){
  
  boot.coef <- rbinom(days, 1, shootings/days)

  if((length(which(boot.coef == 1)) == 0) || (length(which(boot.coef == 1)) == 1) || (length(which(boot.coef == 1)) == 2)){
    return(NA)
  } else {
    
  tau.boot <- diff(which(boot.coef == 1))-1

  tau.boot.sd <- sd(tau.boot)
  tau.boot.mean <- mean(tau.boot)
  tau.boot.variation <- (tau.boot.sd/tau.boot.mean)
  
  return(tau.boot.variation)
  }
}

bootstrap.rep <- replicate(10000, bootstrap.coef.var(365, 10)) # 2019 results 
```


## 2019 Bootstrap Histogram (Centered at 0) 

```{r}
hist(bootstrap.rep-1, breaks="FD", main ="2019 Bootstrap Histogram")
abline(v=coef.variation-1, col="blue")
abline(v=-coef.variation+1, col="red")
```


## Function for the P-value with 2019 Results 

```{r}
#' computes the P-value for a given year
#'
#' @param days number of days in a year 
#' @param shootings number of shooting that given year 
#' @param coef.variation the coefficient variation for that given year 
#'
#' @return the p-value 


bootstrap.pval<-function(days,shootings,coef.variation){
  
  bootstrap.rep<-replicate(10000, bootstrap.coef.var(days,shootings))
  
  right.tail <- bootstrap.rep-1 >= abs(coef.variation-1) # true=values to the right of the hist 
  left.tail <- bootstrap.rep-1 <= -abs(coef.variation-1) # true=values to the left of the hist 
  right.tail | left.tail # TRUE where they're in the right tail OR left tail
  
  pval<-(sum(right.tail | left.tail, na.rm=TRUE))/sum(!is.na(bootstrap.rep)) #p value 

  return(pval)
}

bootstrap.pval(365,10,coef.variation=coef.variation) # 2019 results 
```


2015 Results 


## Calculating One Coefficient of Variation For The Year 2015

```{r}
tau.2015 <- as.numeric(-diff(shootings.dataset$date[41:47])) #Calculating the time between two successive mass shootings in the year 2015

tau.sd<- sd(tau.2015) 
tau.mean <- mean(tau.2015)

coef.variation.2015 <- tau.sd/tau.mean # Calculating the coefficient of variation for 2015

```


## Bootstrap Function with 2015 Results 

```{r}
bootstrap.rep.15 <- replicate(10000, bootstrap.coef.var(365, 7)) # 2015 results 
```


## 2015 Bootstrap Histogram (Centered at 0) 

```{r}
hist(bootstrap.rep.15-1, breaks="FD", main ="2015 Bootstrap Histogram")
abline(v=coef.variation.2015-1, col="purple")
abline(v=-coef.variation.2015+1, col="orange")
```


## P-value with 2015  

```{r}
bootstrap.pval(365,7,coef.variation=coef.variation.2015)
```


# Question 7

## Total Victims Graphs 

```{r}
a <- shootings.dataset[86:118, 3] 

d <- shootings.dataset[86:118, 7]

x <- shootings.dataset[1:85, 3]

w <- shootings.dataset[1:85, 7]

par(mfrow=c(2,2))

plot(a, d, main = "Mass Shootings Total Victims Before 9/11/2001",xlab = "Year", ylab = "Total Victims", col="red")
plot(x, w, main = "Mass Shootings Total Victims  After 9/11/2001", xlab = "Year", ylab = "Total Victims", ylim = c(0, 60), col="blue")
```


## Injured Graphs 

```{r}
a<- shootings.dataset[86:118, 3] 

c<- shootings.dataset[86:118, 6]

x <- shootings.dataset[1:85, 3]

y<- shootings.dataset[1:85, 6]

par(mfrow=c(2,2))

plot(a, c, main = "Mass Shootings Injured Before 9/11/2001",xlab = "Year", ylab = "Injured", ylim = c(0, 60), col="purple")
plot(x, y, main = "Mass Shootings Injured After 9/11/2001", xlab = "Year", ylab = "Injured",ylim = c(0, 60), col="orange")

```


## All Four Graphs Combined

```{r}
par(mfrow=c(2,2))

plot(a, d, main = "Total Victims Before 9/11/2001",xlab = "Year", ylab = "Total Victims", col="red")
plot(x, w, main = "Total Victims  After 9/11/2001", xlab = "Year", ylab = "Total Victims", ylim = c(0, 60), col="blue")
plot(a, c, main = "Injured Before 9/11/2001",xlab = "Year", ylab = "Injured", ylim = c(0, 60), col="purple") 
plot(x, y, main = "Injured After 9/11/2001", xlab = "Year", ylab = "Injured",ylim = c(0, 60), col="orange")
```







