---
title: "MA350 Final Project"
authors: Emily Tumbaco, Lorena Mancino, Rob Golebieski
output: html_notebook
---

# Are Mass Shooting Really That Random? 


```{r}
shootings.dataset<-read.csv("data.csv") #importing the dataset 
```


```{r}
# install.packages('dplyr') this package contains the mutate function which we used below 

library(dplyr)

shootings.dataset <- mutate(shootings.dataset, date= as.Date(date, format = "%m/%d/%Y"))
shootings.dataset$date # converting the dates in the data set to actual dates in R
```


### Mass Shootings Before 9/11/2001 

```{r}
a <- shootings.dataset[86:118, 3]

b <- shootings.dataset[86:118, 5]

plot(a, b)
```
This is a graph of all of the mass shootings that occured before 9/11/2001. The x axis contains the dates, and the y axis contains the number of fatalities from each shooting. 


### Mass Shootings After 9/11/2001 

```{r}
x <- shootings.dataset[1:85, 3]

y <- shootings.dataset[1:85, 5]

plot(x, y)
```
This is a graph of all of the mass shootings that occured after 9/11/2001. The x axis contains the dates, and the y axis contains the number of fatalities from each shooting. 


```{r}
# plot both graphs here
```




















