---
title: "MA350 Final Project"
authors: Emily Tumbaco, Lorena Mancino, Rob Golebieski
output: html_notebook
---

# Are Mass Shooting Really That Random? 

## Problem 1 

```{r}
shootings.dataset<-read.csv("data.csv") #importing the dataset 
```


```{r}
# install.packages('dplyr') # this package contains the mutate function which we used below 

library(dplyr)

shootings.dataset <- mutate(shootings.dataset, date= as.Date(date, format = "%m/%d/%Y"))
shootings.dataset$date # converting the dates in the data set to actual dates in R
```


### Mass Shootings Before 9/11/2001 

```{r}
a <- shootings.dataset[86:118, 3]

b <- shootings.dataset[86:118, 5]

plot(a, b, main = "Mass Shootings Before 9/11/2001",xlab = "Year", ylab = "Fatalities")
```
This is a graph of all of the mass shootings that occured before 9/11/2001. The x axis contains the dates, and the y axis contains the number of fatalities from each shooting. 



### Mass Shootings After 9/11/2001 

```{r}
x <- shootings.dataset[1:85, 3]

y <- shootings.dataset[1:85, 5]

plot(x, y,main = "Mass Shootings After 9/11/2001", xlab = "Year", ylab = "Fatalities")
```
This is a graph of all of the mass shootings that occured after 9/11/2001. The x axis contains the dates, and the y axis contains the number of fatalities from each shooting. 


<<<<<<< HEAD
=======
```{r}
par(mfrow=c(2,2))
plot(a, b, main = "Mass Shootings Before 9/11/2001",xlab = "Year", ylab = "Fatalities")
plot(x, y,main = "Mass Shootings After 9/11/2001", xlab = "Year", ylab = "Fatalities")
```


## Problem 4 

### Raster Plot - 2019 

```{r}
from.date <- as.Date('2019-01-01', format = "%Y-%m-%d")
to.date <- as.Date('2019-12-31', format = "%Y-%m-%d")

plot(NA, xlim=c(from.date, to.date), ylim=c(0,10), xlab="Day of Year", ylab="y")

segments(shootings.dataset$date, 0, y1=10)
```
>>>>>>> e5943566c0c074e2495950e81c2312d2dbd6576f


### Raster Plot - 1999 

```{r}
from.date.1 <- as.Date('1999-01-01', format = "%Y-%m-%d")
to.date.1 <- as.Date('1999-12-31', format = "%Y-%m-%d")

plot(NA, xlim=c(from.date.1, to.date.1), ylim=c(0,10), xlab="Day of Year", ylab="y")

segments(shootings.dataset$date, 0, y1=10)
```


### Raster Plot - 2018

```{r}
from.date.2 <- as.Date('2018-01-01', format = "%Y-%m-%d")
to.date.2 <- as.Date('2018-12-31', format = "%Y-%m-%d")

plot(NA, xlim=c(from.date.2, to.date.2), ylim=c(0,10), xlab="Day of Year", ylab="y")

segments(shootings.dataset$date, 0, y1=10)
```

From looking at the raster plots, the mass shootings seem to occur in clusters. Every 3 months, there are approximately 3 shootings clustered together. The highest frequency of shootings occurs in the Fall months, and the lowest frequency occurs in the Summer. 


## Problem 5

```{r}
tau.2019 <- as.numeric(-diff(shootings.dataset$date[2:11]))

tau.sd<- sd(tau.2019)
tau.mean <- mean(tau.2019)

coef.variation <- tau.sd/tau.mean
coef.variation
```


```{r}
rbinom(365, 1, (10/365))

```





